### 2018-02-07 extensions.tex
added missing .tex files

% !TEX root = hazelnut-dynamics.tex
\newcommand{\extensionsSec}{Extensions to Hazelnut Live}
\section{\protect\extensionsSec} % don't like the all-caps thing that the template does, so protecting it from that
\label{sec:extensions}
\subsection{Product Types}

\subsection{Sum Types}

\subsection{Recursive Types}

\subsection{Polymorphism}

\subsection{Mutable References}
\cy{if we add references, would have to talk about how commutativity isn't conserved}