### 2018-02-20 implementation.tex

% !TEX root = hazelnut-dynamics.tex

\newcommand{\implementationSec}{Implementation}
\section{\protect\implementationSec}
\label{sec:implementation}

\rkc{there should be at least a brief section about the implementation, right?}

\rkc{describe the UI that allows a mix of text/structured edits that ensure
well-formed edit states}