### 2018-02-26 fig-complete.tex
metatheory definitions, wip

\begin{figure}[t]
\rkc{TODO}
\caption{Complete H-types, H-expressions, and dynamic H-expressions}
\label{fig:complete}
\end{figure}