### 2018-03-08 examples-1.tex
splitting Section 2 subsections into separate files
to help avoid merge conflicts

\subsection{Live Programming with Incomplete Expressions}

The process of writing new program fragments includes many states in which the
program under construction is, by definition, incomplete.
%
It is natural for a programmer to ``jump back and forth'' between different
places in the program, iteratively and alternately developing the missing
pieces.
%
The primary benefit of our approach is that such programs can be evaluated in
order to provide useful feedback during this editing workflow.

\input{fig-grades-example}

Consider a teacher who is in the midst of developing a \HazelnutLive{} notebook
(depicted in \autoref{fig:grades-example}) to compute final student grades at
the end of a term.
%
In the first cell in \autoref{fig:grades-example}, the teacher defines a record
type, \li{student_data}, for each student's course data.
%
In our simplified example, this includes the student's name, of type
\li{string}, and five grades, each of type \li{float}.

\overviewExample{1}{Computing Weighted Averages}
%
The teacher begins writing a \li{weighted_average} function that will
compute a weighted average, of type \li{float}, for each \li{student_data}
record.
%
The body of \li{weighted_average} is not yet written---marked with the
\emph{empty hole} on line \rkc{XXX}--- but the teacher jumps ahead to map the
function over the \li{grades} list.

\begin{lstlisting}
let weighted_average(g: student_data) =
  ??;

let weighted_averages = List.map weighted_average grades;
\end{lstlisting}

\noindent
%
\HazelnutLive{} evaluates this incomplete program, showing its
\emph{indeterminate} result in the sidebar on the right half of
\autoref{fig:grades-example}; this value is a list whose length is the same as
\li{grades}, where each element is the (indeterminate) result of evaluating the
hole expression inside \li{weighted_average}.
%
Notice that the hole on line \rkc{XXX} is assigned a unique static identifier
(\ie{}~\li{a}) and that the indeterminate results it produces are assigned
corresponding unique dynamic identifiers (\ie{}~\li{a.1}, \li{a.2}, etc.).
%
In contrast, the typical ad-hoc approach to simulating hole expressions, namely,
using placeholder expressions like \li{raise "Not yet implemented."}, would halt
the program as soon it is evaluated (in this case, inside the definition of
\li{List.map} when the function is applied to the first element of \li{grades}.
%
By ``evaluating around'' the hole in \HazelnutLive{}, the programmer can at
least see that the resulting expression does indeed evaluate to a list with the
correct number of (indeterminate) values.

Now the teacher returns to \li{weighted_average} to work on the missing
expression, with the goal of computing a weighted sum of each of a student's
grades.
%
The teacher uses a \HazelnutLive{} edit action to create a sum expression,
filling in the first summand by weighting the \li{hw1} score to account for 10\%
of the weighted average; the rest of the computation is a new hole expression
(\li{??_b}), yet to be filled in.

\begin{lstlisting}
let weighted_average(g: student_data) =
  (10.0 *. g.hw1) +. ??_b;
\end{lstlisting}

\noindent
%
\HazelnutLive{} immediately runs this incomplete program, displaying the updated
list of indeterminate expressions \li{[760.0 + ??_b.1, 880.0 + ??_b.2, ...]}.
%
Right away, the teacher recognizes that the values are too large; they should be
at most \li{100.0}.
%
The teacher realizes that representing percentage points as \li{float}s requires
that the constant on line \rkc{XXX} ought to be \li{0.10} instead.
%
Because of this live feedback, the teacher corrects this error right away and
avoids making similar programming errors in the rest of the calculation.
%
The teacher continues to build the rest of the arithmetic expression until it is
complete (there are no longer any expression holes), and the result of
immediately running the finished program shows that each of values in the final
result list is in the range \li{0.0} to \li{100.0}.

\overviewExample{2}{Assigning Letter Grades}
%
The teacher's next task is to map the weighted averages to the letter grades A
through F (we will consider only ``whole'' letter grades, for simplicity).
%
The \li{grade_cutoffs} record type describes the minimum cutoff for each of
these six possible grades.
%
Initially, each value in the \li{cutoffs} record is a hole.

\begin{lstlisting}
type grade_cutoffs =
  { a: float, b: float, c: float, d: float, f: float };

let cutoffs =
  { a = ??_a, b = ??_b, c = ??_c, d = ??_d, f = ??_f };
\end{lstlisting}

\noindent
%
%% initially all holes, because it will be different year-to-year based on the %
%data, differences in course difficulty, and to satisfy fairness criteria.
%
Before starting to fill in the cutoffs, the teacher jumps ahead to write a
function \li{letter_grade} that will make the connection between \li{cutoffs}
and \li{weighted_averages}.
%
Because she intends to look at the data to help select the cutoff values, the
teacher sorts the \li{weighted_averages} and then maps them to
\li{letter_grades}.

\begin{lstlisting}
let letter_grade(n: weighted_average) =
  if n >= cutoffs.a then "A" else
  if n >= cutoffs.b then "B" else
  if n >= cutoffs.c then "C" else
  if n >= cutoffs.d then "D" else
  if n >= cutoffs.f then "F" else "Incomplete";

let sorted_weighted_averages = List.sort weighted_averages;

let letter_grades = List.map letter_grade sorted_weighted_averages;
\end{lstlisting}

\noindent
%
When \HazelnutLive{} runs this program, the guard of the outermost conditional
(\li{n >= cutoffs.a} on line \rkc{XXX}), is indeterminate because \li{cutoffs.a}
is.
%
Therefore, each of the indeterminate expressions in \li{letter_grades} is the
entire expression body, albeit with different bindings for \li{n}.
%
Displaying such ``large'' indeterminate expressions can quickly consume all
available screen space, overwhelming the user with too much information.
%
\autoref{fig:XXX} shows how \HazelnutLive{} renders large indeterminate
expressions (\ie{}~anything other than ``small'' indeterminate value (a constant
or hole) or list of small values) simply as \li{..} to save space.
%
Hovering over this abbreviation (as shown in \autoref{fig:XXX} displays the full
indeterminate expression---as well as the evaluation environment that surrounds
it---as a tooltip.
%
\autoref{sec:discussion} discusses this and other user interface concerns when
trying to display useful live feedback without overwhelming the user.
%
These usability factors are beyond the scope of our work, which is to define
semantic foundations on which such user interfaces can be built.

To start deciding \li{cutoffs}, the teacher looks at the \li{weighted_averages}
sorted in descending order.
%
The data shows a natural gap between \li{92.2} and \li{89.5}.
%
So, she chooses to use \li{92.0} as the cutoff for A.
%
Resuming the computation from before, \HazelnutLive{} resolves the conditional
expressions for the first \rkc{XXX} indeterminate expressions, because each of
those \li{n} values was greater than \li{cutoffs.a}.
%
The remaining \rkc{XXX} expressions also proceeded to evaluate the first guard,
and are now indeterminate at the guard for the second conditional.

Before assigning other cutoff values, the teacher would like to get a sense of
whether this first choice is a good one.
%
She jumps ahead to to write a function that computes the distribution of letter
grades.

\begin{lstlisting}
let compute_distribution(list: list(float)) =
  let n = List.length letter_grades in
  List.map
    (\x -> (x, showPercentage (List.length (List.filter ((==) x) list)) /. n))
    ["A","B","C","D","F","Incomplete"];

let distribution = compute_distribution(letter_grades);
\end{lstlisting}

\noindent
%
Running \li{compute_distribution} shows that the percentage of As is
\rkc{XXX\%}, which is smaller than what the teacher would like;
the remaining percentages are all indeterminate.
%
Returning to the value of \li{sorted_data}, she sees a cluster around \li{89.0}
and then another gap between \li{88.2} and {85.5}.
%
So, the teacher adjusts \li{cutoffs.a} to be \li{88.0}.
%
Because this edit is outside of a hole expression, \HazelnutLive{} discards the
previous execution state and reevaluates the entire program.
%
Existing techniques for incremental computation~\cite{XXX,XXX}, however, could
be applied to seek opportunities for reuse even when non-hole expressions are
modified.
%
Because the focus of work is on the novel implications of running programs with
holes, our calculus and implementation supports caching ``edit-and-resume'' only
for the novel situation in which evaluation proceeds around holes.
%
After re-evaluation, the percentage of As becomes \rkc{XXX\%}, which better
matches the teacher's intention.

In this fashion, the teacher continues down the list of sorted averages,
determining appropriate values for each cutoff.
%
Whenever the teacher is only filling in the ``next'' cutoff, the computation
from before can simply be resumed.
%
Overall, throughout the workflow described in these two examples, the programmer
can continue to evaluate the program, and receive meaningful feedback, while
going back and forth between different pieces under development.