### 2018-03-28 appendix-defns.tex
draft of sec 3.0 through 3.1

% !TEX root = hazelnut-dynamics.tex

\newcommand{\additionalDefnsSec}{Additional Definitions for Hazelnut Live}
\section{\protect\additionalDefnsSec} % don't like the all-caps thing that the template does, so protecting it from that
\label{sec:additional-defns}

\input{fig-complete}