### 2018-04-01 appendix-defns.tex
finished draft of 3.2

% !TEX root = hazelnut-dynamics.tex

\newcommand{\additionalDefnsSec}{Additional Definitions for Hazelnut Live}
\section{\protect\additionalDefnsSec} % don't like the all-caps thing that the template does, so protecting it from that
\label{sec:additional-defns}

% No weakening for Gammas in Delta: 
% If $\hasType{\Delta, \Dbinding{u}{\hGamma}{\tau}}{\hGamma'}{d}{\tau'}$ then $\hasType{\Delta, \Dbinding{u}{\hGamma, x : \tau''}{\tau}}{\hGamma'}{d}{\tau'}$.

\input{fig-complete}