### 2018-04-17 extensions.tex

% !TEX root = hazelnut-dynamics.tex
\newcommand{\extensionsSec}{Extensions to Hazelnut Live}
\section{\protect\extensionsSec} % don't like the all-caps thing that the template does, so protecting it from that
\label{sec:extensions}

\subsection{Numbers}
\include{numbers} %(from Hazelnut)

%\subsection{Product Types}
%\include{prodtypes}

\subsection{Sum Types}
\include{sumtypes}

%\subsection{Recursive Types}
%\subsection{Polymorphism}
%\subsection{Mutable References}
%\cy{if we add references, would have to talk about how commutativity isn't conserved}