### 2018-04-17 appendix-defns.tex
submission

% !TEX root = hazelnut-dynamics.tex

\newcommand{\additionalDefnsSec}{Additional Definitions for Hazelnut Live}
\section{\protect\additionalDefnsSec} % don't like the all-caps thing that the template does, so protecting it from that
\label{sec:additional-defns}

\subsection{Substitution}
\label{sec:substitution}
\judgbox
  {[\dexp/x]\dexp' = \dexp''}
  {$\dexp''$ is obtained by substituting $\dexp$ for $x$ in $\dexp'$}
  %% {$\dexp''$ is the result of substituting $\dexp$ for $u$ in $\dexp'$}

\vspace{5pt}
\judgbox
  {[\dexp/x]\sigma = \sigma'}
  {$\sigma'$ is obtained by substituting $\dexp$ for $x$ in $\sigma$}
  %% {$\dexp''$ is the result of substituting $\dexp$ for $u$ in $\dexp'$}
\[
\begin{array}{lcll}
\substitute{\dexp}{x}{c}
&=&
c\\
\substitute{\dexp}{x}{x}
&=&
\dexp
\\%
\substitute{\dexp}{x}{y}
&=&
y
& \text{when $x \neq y$}
\\
\substitute{\dexp}{x}{\halam{y}{\htau}{\dexp'}}
&=&
\halam{y}{\htau}{\substitute{\dexp}{x}{\dexp'}}
& \text{when $x \neq y$ and $y \notin \fvof{d}$}
\\
\substitute{\dexp}{x}{\dap{\dexp_1}{\dexp_2}}
&=&
\dap{(\substitute{\dexp}{x}{\dexp_1})}{\substitute{\dexp}{x}{\dexp_2}}
\\
\substitute{\dexp}{x}{\dehole{u}{\subst}{}}
&=&
\dehole{u}{\substitute{\dexp}{x}{\subst}}{}
\\
\substitute{\dexp}{x}{\dhole{\dexp'}{u}{\subst}{}}
&=&
\dhole{\substitute{\dexp}{x}{\dexp'}}{u}{\substitute{\dexp}{x}{\subst}}{}
\\
\substitute{\dexp}{x}{\dcasttwo{\dexp'}{\htau_1}{\htau_2}}
&=&
\dcasttwo{(\substitute{\dexp}{x}{\dexp'})}{\htau_1}{\htau_2}
\\
\substitute{\dexp}{x}{\dcastfail{\dexp'}{\htau_1}{\htau_2}}
&=&
\dcastfail{(\substitute{\dexp}{x}{\dexp'})}{\htau_1}{\htau_2}
\\
\substitute{\dexp}{x}{\cdot}
&=&
\cdot\\
\substitute{\dexp}{x}{\sigma, d/y}
&=&
\substitute{\dexp}{x}{\sigma}, \substitute{\dexp}{x}{d}/y
%% {[\dexp_1 / x] \dcast{\htau}{\dexp}}
%% &=&
%% \dcast{\htau}{[\dexp_1 / x] \dexp}
\end{array}
\]

\vspace{5pt}
\begin{lem}[Substitution] \label{thm:substitution}~
\begin{enumerate}[nolistsep]
\item If $\hasType{\hDelta}{\hGamma, x : \htau'}{d}{\htau}$ and $\hasType{\hDelta}{\hGamma}{d'}{\htau'}$ then $\hasType{\hDelta}{\hGamma}{[d'/x]d}{\htau}$.
\item If $\hasType{\hDelta}{\hGamma, x : \htau'}{\sigma}{\hGamma'}$ and $\hasType{\hDelta}{\hGamma}{d'}{\htau'}$ then $\hasType{\hDelta}{\hGamma}{[d'/x]\sigma}{\hGamma'}$.
\end{enumerate}
\end{lem}
\begin{proof} By rule induction on the first assumption in each case. The conclusion follows from the definition of substitution in each case. \end{proof}

\subsection{Canonical Forms} 
\begin{lem}[Canonical Value Forms]\label{thm:canonincal-value-forms}
  If $\hasType{\hDelta}{\emptyset}{\dexp}{\htau}$ and $\isValue{\dexp}$
  then $\htau\neq\tehole$ and
  \begin{enumerate}[nolistsep]
    \item If $\htau=b$ then $\dexp=c$.
    \item If $\htau=\tarr{\htau_1}{\htau_2}$
          then $\dexp=\halam{x}{\htau_1}{\dexp'}$
          where $\hasType{\hDelta}{x : \htau_1}{\dexp'}{\htau_2}$.
  \end{enumerate}
\end{lem}

\begin{lem}[Canonical Boxed Forms]\label{thm:canonical-boxed-forms}
  If $\hasType{\hDelta}{\emptyset}{\dexp}{\htau}$ and $\isBoxedValue{\dexp}$
  then
  \begin{enumerate}[nolistsep]
    \item If $\htau=b$ then $\dexp=c$.
    \item If $\htau=\tarr{\htau_1}{\htau_2}$ then either
      \begin{enumerate}
        \item[i.]
          $\dexp=\halam{x}{\htau_1}{\dexp'}$
          where $\hasType{\hDelta}{x : \htau_1}{\dexp'}{\htau_2}$, or
        \item[ii.]
          $\dexp=\dcasttwo{\dexp'}{\tarr{\htau_1'}{\htau_2'}}{\tarr{\htau_1}{\htau_2}}$
          where $\tarr{\htau_1'}{\htau_2'}\neq\tarr{\htau_1}{\htau_2}$
          and $\hasType{\hDelta}{\emptyset}{\dexp'}{\tarr{\htau_1'}{\htau_2'}}$.
      \end{enumerate}
    \item If $\htau=\tehole$
          then $\dexp=\dcasttwo{\dexp'}{\htau'}{\tehole}$
          where $\isGround{\htau'}$
          and $\hasType{\hDelta}{\emptyset}{\dexp'}{\htau'}$.
  \end{enumerate}
\end{lem}

\begin{lem}[Canonical Indeterminate Forms]
  If $\hasType{\hDelta}{\emptyset}{\dexp}{\htau}$
  and $\isIndet{\dexp}$
  then either
  \begin{enumerate}[nolistsep]
    \item
      $\dexp=\dehole{u}{\subst}{}$
      and $\Dbinding{u}{\Gamma'}{\htau}\in\hDelta$, or
    \item
      $\dexp=\dhole{\dexp'}{u}{\subst}{}$
      and $\isFinal{\dexp'}$
      and $\hasType{\hDelta}{\emptyset}{\dexp'}{\htau'}$
      and $\Dbinding{u}{\Gamma'}{\htau}\in\hDelta$, or
    \item
      $\dexp=\dap{\dexp_1}{\dexp_2}$
      and $\hasType{\hDelta}{\emptyset}{\dexp_1}{\tarr{\htau_2}{\htau}}$
      and $\hasType{\hDelta}{\emptyset}{\dexp_2}{\htau_2}$
      and $\isIndet{\dexp_1}$
      and $\isFinal{\dexp_2}$
      and $\dexp_1\neq\dcasttwo{\dexp_1}{\tarr{\htau_3}{\htau_4}}
                                        {\tarr{\htau_3'}{\htau_4'}}$, or
    %% \item
    %%   \begin{enumerate}
    %%     \item blah
    %%     \item blah
    %%     \item blah
    %%   \end{enumerate}
    \item
      $\htau=b$
      and $\dexp=\dcasttwo{\dexp'}{\tehole}{b}$
      and $\isIndet{\dexp'}$
      and $\dexp'\neq\dcasttwo{\dexp''}{\htau'}{\tehole}$, or
    \item
      $\htau=b$
      and $\dexp=\dcastfail{\dexp'}{\htau'}{b}$
      and $\isGround{\htau'}$
      and $\htau'\neq{b}$
      and $\hasType{\hDelta}{\emptyset}{\dexp'}{\htau'}$, or
    \item
      $\htau=\tarr{\htau_{11}}{\htau_{12}}$
      and $\dexp=\dcasttwo{\dexp'}{\tarr{\htau_1}{\htau_2}}
                                  {\tarr{\htau_{11}}{\htau_{12}}}$
      and $\isIndet{\dexp'}$
      and $\tarr{\htau_1}{\htau_2}\neq\tarr{\htau_{11}}{\htau_{12}}$, or
    \item
      $\htau=\tarr{\tehole}{\tehole}$
      %% $\htau=\tarr{\htau_{11}}{\htau_{12}}$
      %% and $\htau_{11}=\tehole$
      %% and $\htau_{12}=\tehole$
      and $\dexp=\dcastthree{\dexp'}{\tehole}{\tehole}{\tehole}$
      and $\isIndet{\dexp'}$
      and $\dexp'\neq\dcasttwo{\dexp''}{\htau'}{\tehole}$, or
    \item
      $\htau=\tarr{\tehole}{\tehole}$
      %% $\htau=\tarr{\htau_{11}}{\htau_{12}}$
      %% and $\htau_{11}=\tehole$
      %% and $\htau_{12}=\tehole$
      and $\dexp=\dcastfail{\dexp'}{\htau'}{\tarr{\tehole}{\tehole}}$
      %% and $\dexp=\dcastfail{\dexp'}{\htau'}{\tarr{\htau_{11}}{\htau_{12}}}$
      and $\htau'\neq\htau$
      and $\isGround{\htau'}$
      and $\isIndet{\dexp'}$
      and $\hasType{\hDelta}{\emptyset}{\dexp'}{\htau'}$, or
    \item
      $\htau=\tehole$
      and $\dexp=\dcasttwo{\dexp'}{\htau'}{\tehole}$
      and $\isGround{\htau'}$
      and $\isIndet{\dexp'}$.
  \end{enumerate}
\end{lem}

The proofs for all three of these theorems follow by straightforward rule induction.

% No weakening for Gammas in Delta: 
% If $\hasType{\Delta, \Dbinding{u}{\hGamma}{\tau}}{\hGamma'}{d}{\tau'}$ then $\hasType{\Delta, \Dbinding{u}{\hGamma, x : \tau''}{\tau}}{\hGamma'}{d}{\tau'}$.

\clearpage
\subsection{Complete Programs}

\input{fig-complete}

\subsection{Multiple Steps}
\input{fig-multi-step}

\clearpage
\subsection{Hole Filling}\label{sec:hole-filling}
\begin{lem}[Filling] ~
  \begin{enumerate}[nolistsep]
  \item If $\hasType{\hDelta, \Dbinding{u}{\hGamma'}{\htau'}}{\hGamma}{\dexp}{\tau}$
  and $\hasType{\hDelta}{\hGamma'}{\dexp'}{\htau'}$
  then $\hasType{\hDelta}{\hGamma}{\instantiate{\dexp'}{u}{\dexp}}{\tau}$.
  \item If $\hasType{\hDelta, \Dbinding{u}{\hGamma'}{\htau'}}{\hGamma}{\sigma}{\hGamma''}$
  and $\hasType{\hDelta}{\hGamma'}{\dexp'}{\htau'}$
  then $\hasType{\hDelta}{\hGamma}{\instantiate{\dexp'}{u}{\sigma}}{\hGamma''}$.
  \end{enumerate}
\end{lem}
\begin{proof}
In each case, we proceed by rule induction on the first assumption, appealing to the Substitution Lemma as necessary.
\end{proof}

We need the following auxiliary definitions, which lift hole filling to evaluation contexts taking care to consider the special situation where the mark is inside the hole that is being filled, to prove the Commutativity theorem.
\input{fig-evalctx-instantiation}
\begin{lem}[Substitution Commutativity]
  If 
  \begin{enumerate}[nolistsep]
  	\item $\hasType{\hDelta, \Dbinding{u}{\hGamma'}{\htau'}}{x : \htau_2}{\dexp_1}{\tau}$ and 
  	\item $\hasType{\hDelta, \Dbinding{u}{\hGamma'}{\htau'}}{\emptyset}{\dexp_2}{\htau_2}$ and 
  	\item $\hasType{\hDelta}{\hGamma'}{\dexp'}{\htau'}$
  \end{enumerate}

  then  $\instantiate{d'}{u}{\substitute{d_2}{x}{d_1}} = \substitute{\instantiate{d'}{u}{d_2}}{x}{\instantiate{d'}{u}{d_1}}$.
\end{lem}
\begin{proof}
We proceed by structural induction on $d_1$ and rule induction on the typing premises, which serve to ensure that the free
variables in $d'$ are accounted for by every closure for $u$.
\end{proof}

\begin{lem}[Instruction Commutativity]
  If 
  \begin{enumerate}[nolistsep]
  	\item $\hasType{\hDelta, \Dbinding{u}{\hGamma'}{\htau'}}{\emptyset}{\dexp_1}{\tau}$ and 
  	\item $\hasType{\hDelta}{\hGamma'}{\dexp'}{\htau'}$ and 
  	\item $\reducesE{}{\dexp_1}{\dexp_2}$
  \end{enumerate}

  then $\reducesE{}{\instantiate{\dexp'}{u}{\dexp_1}}
                     {\instantiate{\dexp'}{u}{\dexp_2}}$.
\end{lem}
\begin{proof}
We proceed by cases on the instruction transition assumption (no induction is needed). For Rule \rulename{ITBeta}, we defer to the Substitution Commutativity lemma above. For the remaining cases, the conclusion follows from the definition of hole filling.
\end{proof}

\begin{lem}[Filling Totality]
Either $\inhole{u}{\evalctx}$ or $\instantiate{d}{u}{\evalctx}=\evalctx'$ for some $\evalctx'$.
\end{lem}
\begin{proof} We proceed by structural induction on $\evalctx$. Every case is handled by one of the two judgements. \end{proof}

\begin{lem}[Discarding] If 
	\begin{enumerate}[nolistsep]
	\item $\selectEvalCtx{d_1}{\evalctx}{\dexp_1'}$ and 
	\item $\selectEvalCtx{d_2}{\evalctx}{\dexp_2'}$ and 
	\item $\inhole{u}{\evalctx}$
	\end{enumerate}

	then $\instantiate{d}{u}{d_1} = \instantiate{d}{u}{d_2}$.
\end{lem}
\begin{proof} We proceed by structural induction on $\evalctx$ and rule induction on all three assumptions. Each case follows from the definition of instruction selection and hole filling. \end{proof}

\begin{lem}[Filling Distribution] If 
	$\selectEvalCtx{d_1}{\evalctx}{d_1'}$ and $\instantiate{d}{u}{\evalctx}=\evalctx'$ then $\selectEvalCtx{\instantiate{d}{u}{d_1}}{\evalctx'}{\instantiate{d}{u}{d_1'}}$.
\end{lem}
\begin{proof} We proceed by rule induction on both assumptions. Each case follows from the definition of instruction selection and hole filling. \end{proof}

\begin{thm}[Commutativity]
  If 
  \begin{enumerate}[nolistsep]
  \item $\hasType{\hDelta, \Dbinding{u}{\hGamma'}{\htau'}}{\emptyset}{\dexp_1}{\tau}$ and 
  \item $\hasType{\hDelta}{\hGamma'}{\dexp'}{\htau'}$ and 
  \item $\multiStepsTo{\dexp_1}{\dexp_2}$
\end{enumerate}

  then $\multiStepsTo{\instantiate{\dexp'}{u}{\dexp_1}}
                     {\instantiate{\dexp'}{u}{\dexp_2}}$.
\end{thm}
\begin{proof}
By rule induction on assumption (3). The reflexive case is immediate. In the inductive case, we proceed by rule induction on the stepping premise. There is one rule, Rule~\rulename{Step}. By Filling Totality, either $\inhole{u}{\evalctx}$ or $\instantiate{d}{u}{\evalctx} = \evalctx'$. In the former case, by Discarding, we can conclude by \rulename{MultiStepRefl}. In the latter case, by Instruction Commutativity and Filling Distribution we can take a \rulename{Step}, and we can conclude via \rulename{MultiStepSteps} by applying Filling, Preservation and then the induction hypothesis.
\end{proof}

We exclude these proofs and definitions from the Agda mechanization
for two reasons.
%
First, fill-and-resume is merely an optimization, and unlike the meta
theory of \Secref{sec:calculus}, these properties are generally not
conserved by certain reasonable extensions of the core
calculus~(e.g., reference cells and other non-commuting effects).
%
Second, to properly encode the hole filling operation, such a
mechanization requires a significantly more complex representation of
hole environments; unfortunately, Agda cannot be easily convinced that
the definition of hole filling is well-founded (\citet{Nanevski2008}
establish that it is in fact well-founded).
%
By contrast, the developments in \Secref{sec:calculus} do not require
these more complex (and somewhat problematic) representations.

\subsection{Confluence, and Friends}\label{sec:confluence}
There are various ways to encode the intuition that ``evaluation order does not matter''. One way to do so is
by establishing a confluence property (which is closely related to
the Church-Rosser property \cite{church1936some}).

The most general confluence property does not hold for the dynamic
semantics in Sec.~\ref{sec:calculus} for the usual reason: We do not
reduce under binders (\citet{DBLP:conf/birthday/BlancLM05} discuss the
standard counterexample).
%
We could recover confluence by specifying reduction under binders,
either generally or in a more restricted form where only closed
sub-expressions are
reduced \cite{DBLP:journals/tcs/CagmanH98,DBLP:conf/birthday/BlancLM05,levy1999explicit}.
%
However, reduction under binders conflicts with the standard implementation approaches
for most programming languages \cite{DBLP:conf/birthday/BlancLM05}.
%
A more satisfying approach considers confluence modulo equality \cite{Huet:1980ng}. 
%
The simplest such approach restricts our interest to top-level expressions
of base type that result in values, in which case the following
special case of confluence does hold (trivially when the only base
type has a single value, but also more generally for other base
types).
\begin{lem}[Base Confluence]
  If $\hasType{\Delta}{\emptyset}{\dexp}{b}$ and 
  $\multiStepsTo{\dexp}{\dexp_1}$
  and $\isValue{\dexp_1}$
  and $\multiStepsTo{\dexp}{\dexp_2}$
  then $\multiStepsTo{\dexp_2}{\dexp_1}$.
\end{lem}
We can then prove the following property, which establishes that fill-and-resume is sound.
\begin{thm}[Resumption]
  If $\hasType{\hDelta, \Dbinding{u}{\hGamma'}{\htau'}}{\emptyset}{\dexp}{b}$
  and $\hasType{\hDelta}{\hGamma'}{\dexp'}{\htau'}$ 
  and $\multiStepsTo{\dexp}{\dexp_1}$
  and $\multiStepsTo{\instantiate{\dexp'}{u}{\dexp}}{\dexp_2}$
  and $\isValue{\dexp_2}$
  then $\multiStepsTo{\instantiate{\dexp'}{u}{\dexp_1}}{\dexp_2}$.
  \begin{proof}
    By Commutativity,
    $\multiStepsTo{\instantiate{\dexp'}{u}{\dexp}}
                  {\instantiate{\dexp'}{u}{\dexp_1}}$.
    By Base Confluence, we can conclude.
  \end{proof}
\end{thm}