### 2018-04-17 discussion.tex

% !TEX root = hazelnut-dynamics.tex
\newcommand{\discussionSection}{Conclusion}
\section{\protect\discussionSection} % don't like the all-caps thing that the template does, so protecting it from that
\label{sec:discussion}
To conclude, we quote Weinberg from The Psychology of Computer Programming (1998): ``how truly sad it is that just at the very moment
when the computer has something important to tell us, it starts
speaking gibberish.''

% \todo{discussion}

% \rkc{something about UI choices for displaying large indeterminate expressions}

% \rkc{something about UI choices for displaying hole environment traces}

% \rkc{something about patterns of use for debugging, such as inserting holes
% around all expressions in tail position}

% \rkc{something about visualizing hole environment traces}

% \rkc{\citet{Seidel2016} define a partial evaluation with a form of holes. compare.
% compare ``flattened'' sumList indeterminate result to their trace. could be
% useful to study whether evaluating the type-correct parts of the evaluation
% trace help students focus on the part that has the error.}

% \rkc{quick mention that lazy evaluation with run-time exceptions for holes is
% not enough. this would delay the failure a bit longer, but not indefinitely.}

% live mobile application development systems like Flutter \cite{flutter};