### 2018-04-30 extensions.tex
substantially expanded the related work

% !TEX root = hazelnut-dynamics.tex
\clearpage
\newcommand{\extensionsSec}{Extensions to Hazelnut Live}
\section{\protect\extensionsSec} % don't like the all-caps thing that the template does, so protecting it from that
\label{sec:extensions}

We give two extensions here, numbers and sum types, to maintain parity with \Hazelnut as specified by \citet{popl-paper}.

It is worth observing that these extensions do not make explicit mention of expression holes. The ``non-obvious'' machinery is almost entirely related to casts. Fortunately, there has been excellent work of late on generating ``gradualized'' specifications from standard specifications \cite{DBLP:conf/popl/CiminiS16,DBLP:conf/popl/CiminiS17}. The extensions below, and other extensions of interest, closely follow the output of the gradualizer: \url{http://cimini.info/gradualizerDynamicSemantics/} (which, like our work, is based on the refined account of gradual typing by \cite{DBLP:conf/snapl/SiekVCB15}). The rules for indeterminate forms are mainly where the gradualizer is not sufficient. We leave to future work the related question of automatically generating a hole-aware static and dynamic semantics from a standard language specification.

\include{numbers} %(from Hazelnut)

%\subsection{Product Types}
%\include{prodtypes}

\include{sumtypes}

%\subsection{Recursive Types}
%\subsection{Polymorphism}
%\subsection{Mutable References}
%\cy{if we add references, would have to talk about how commutativity isn't conserved}