### 2018-06-07 rebuttal.txt
edits to rebuttal C

Thank you for your thoughtful comments.

# Review A

> I'm still confused about which version of the language is implemented.
> Are the screenshots in section 2 real, or only the ones in the appendix?

The submitted implementation supports the language of Section 3 plus the
extensions in the appendix. The screenshots in the Appendix show the 
implementation at the time of submission.

All live programming features described in Section 2 are implemented. However,
the figures in Section 2 do depict certain language conveniences that were not 
implemented at the time of submission (per the intro of Section 2). To create 
the figures, we first ran simpler versions of the examples (using encodings 
based on the core calculus), and then we manually post-processed the resulting
HTML/CSS of the editor to depict standard conveniences common to languages like 
Elm and ML, e.g. records, basic pattern matching and syntactic sugar for 
function definitions and lists.

For artifact evaluation, we plan to submit a more recent snapshot of the Hazel
implementation that either substantially narrows or eliminates this 
"convenience gap".

> The HazelnutLive calculus in the paper is really basic, it doesn't even have 
> fixpoint.

Type holes actually allow you to express general recursion (e.g. via the Y 
combinator). See Siek and Taha [2006] for more on embedding the untyped lambda
calculus into a gradually typed lambda calculus like Hazelnut Live.

That said, we agree that an explicit fix operator might be more clear; there is
nothing that would complicate its inclusion, so if requested in our revision, we
would be happy to include it either in the main calculus or as an extension.

# Review C

> While, the technical development is sound, it is mostly based prior work 
> (Hazelnut type system), or a variation on gradually typed calculi, and 
> not very surprising. In fact, evaluation of expression with holes is akin 
> to evaluation of expressions with free variables extended with type casts 
> from gradual typing.

Our focus is the language from the Hazelnut paper, but that paper did not define
a dynamic semantics at all. The solution we propose in this paper certainly
builds on prior work (specifically, gradually typed calculi and CMTT), but we
believe the differences are noteworthy.

1. Gradually typed lambda calculi consider only type holes, and cast failure
immediately aborts evaluation. We do not abort immediately (Sec. 2.4). Moreover,
we consider not just type holes but also expression holes, which operate not
merely as free variables, as suggested, but as closures. Closure tracking is
critical to the user experience in Sec. 2 and to the fill-and-resume feature
described in Sec. 4.

2. Closure tracking has a conceptual relationship with prior work on CMTT
(a non-obvious insight, we believe!), but that work also did not define an
operational semantics nor consider the commutativity property in Sec. 4.

Combining these pieces of prior work with the necessary technical infrastructure
(e.g. indeterminate forms) to be able to prove preservation, progress and
commutativity theorems was not (for us, at least) completely straightforward; we
iterated through several designs before settling on the proposal in the paper.
We are not aware of any comparable related work.

> the calculus lacks recursion and control-flow. Dealing with these efficiently 
> and providing a good user interface for inspecting the resulting reified values 
> is the hard problem in making such an environment.

## Semantics for Recursion and Control Flow

Please see the discussion above regarding recursion.

The sum types extension in the appendix adds primitive control flow. With the
additional page budget, we will include this extension in the main body of the
second phase submission.

## UI for Recursion and Control Flow

We describe our initial UI design considerations for recursion in Sec. 2.2. The
UI allows the user to explore the recursive structure of hole closures
iteratively as described in that section, rather than displaying nested hole
closures all at once, as is the case in the notation used for the core calculus.

Regarding control flow, the UI displays a "breadcrumb" path from the result to
reflect the exploration that the user has done (the basic UI can be seen at the
bottom of Fig. 17). We would be happy to include this more prominently in the
second phase submission (it requires showing a second, more deeply nested hole
environment, so we will need the additional space available in the second phase
submission).

## Effects and Mutation

TODO