### 2018-06-07 rebuttal.txt
more on UI stuff -- just need to talk about effects now

Thank you for your thoughtful feedback!

# Review A

> it doesn't even have fixpoint

Type holes actually allow you to express general recursion (e.g. via the Y 
combinator) - see footnote 9 and [Siek and Taha, 2006] on embedding the 
untyped lambda calculus.

That said, we agree that an explicit fix operator might be more clear; there is
nothing that would complicate its inclusion, so if requested, we would be happy
to include it either in the main calculus or as an extension.

> I'm still confused about which version of the language is implemented.
> Are the screenshots in section 2 real, or only the ones in the appendix?

The screenshots in the Appendix show the submitted implementation, which 
supports all of the live programming features described in Section 2 for 
the language of Section 3 plus the extensions in Appendix B.

For the sake of exposition, the figures in Section 2 do depict certain language
features that were not implemented at submission (as acknowledged in the 
intro to Section 2). To create these figures, we first ran simpler variants of 
the examples (using typical lambda encoding tricks) in our implementation, and 
then we manually edited the resulting HTML/CSS only to add in conveniences 
common to languages like ML/Elm, namely record labels, basic pattern matching 
and syntactic sugar for function definitions and lists.

For artifact evaluation, we plan to submit a more recent snapshot of the Hazel
implementation that either substantially narrows or eliminates this small 
"convenience gap".

# Review C

> While, the technical development is sound, it is mostly based prior work 
> (Hazelnut type system), or a variation on gradually typed calculi, and 
> not very surprising. In fact, evaluation of expression with holes is akin 
> to evaluation of expressions with free variables extended with type casts

Only the external type system comes from the Hazelnut paper (reviewed in 3.1). 
That paper did not define a dynamic semantics at all.

Unlike gradually typed calculi, we do not abort immediately on cast failure 
(Sec. 2.4). Moreover, we consider not only type holes but also expression 
holes, which operate not merely as free variables (which, as suggested, would
be rather uninteresting), but as metavariable closures. Closure tracking is 
critical to the user experience throughout Sec. 2 and to the fill-and-resume 
feature in Sec. 4.

Closure tracking has a relationship with the prior work on CMTT (a non-obvious 
and satisfying logical connection, we believe!), but that work also did not 
define a dynamic semantics, nor did it consider non-empty holes nor the 
commutativity property in Sec. 4.

We carefully combined mechanisms from these prior systems and defined new 
technical infrastructure (e.g. indeterminate forms) to be able to prove clean 
preservation, progress and commutativity theorems mechanically. We are not 
aware of a dynamic semantics with comparable features and metatheory.

> the calculus lacks recursion and control-flow

Please see the response above regarding recursion. The sum types extension 
in Appendix B adds primitive control flow. With the additional page budget, 
we could move this extension into the main body.

> providing a good user interface for inspecting the resulting reified values 
> is the hard problem in making such an environment

We agree that both theoretical and UI considerations are worthy of substantial 
investigation. This paper focused on laying detailed type-theoretic foundations
motivated by one proof-of-concept UI design and several small but non-trivial 
examples.

This UI was designed explicitly with recursion in mind in that it allows 
the user to explore the recursive structure of hole closures iteratively (see 
Sec 2.2). It does not display nested hole closures all at once, as in the 
notation used in the core calculus. We say a little more below the fold.

Branches blocked on a hole can certainly "blow up" in size. A simple, partial 
solution is to visually collapse such sub-expressions by default. Because we 
define a small step semantics, another approach is to pause evaluation when 
this situation occurs and continue only if explicitly requested. Because we 
leave evaluation order undefined, the user could also explicitly shift the 
"evaluation cursor" elsewhere in such a situation.

## Effects and Mutation

TODO

-------

Regarding , the UI displays a "breadcrumb" path from the result to
reflect the exploration that the user has done (the basic UI can be seen at the
bottom of Fig. 17). We would be happy to include this more prominently in the
second phase submission (it requires showing a second, more deeply nested hole
environment, so we will need the additional space available in the second phase
submission).