### 2018-06-08 rebuttal.txt
exactly 600 words before the fold

Thank you for your thoughtful feedback!

# Review A

> it doesn't even have fixpoint

Type holes actually give you general recursion (e.g. via the Y 
combinator) - see footnote 9 and [Siek and Taha, 2006] on embedding the 
untyped lambda calculus.

That said, we agree that an explicit fix operator might be more clear; if 
requested, we would be happy to make this small + simple addition.

> Are the screenshots in section 2 real, or only the ones in the appendix?

The appendix shows the submitted implementation, which 
supports all of the live programming features in Section 2 for 
the language of Section 3 plus Appendix B.

For the sake of exposition, the figures in Section 2 do go beyond the language
of Section 3 (acknowledged in the intro to Section 2). To create these figures, 
we ran simpler variants of the examples (using typical lambda encodings), then 
manually edited the HTML/CSS only to add conveniences from Elm, namely record 
labels, pattern matching and sugar for functions and lists.

(We plan to submit a more recent snapshot of Hazel that substantially narrows 
or eliminates this small "convenience gap" for artifact evaluation.)

# Review C

> mostly based prior work (Hazelnut type system), or a variation on gradually 
> typed calculi, and not very surprising. In fact, evaluation of expression 
> with holes is akin to evaluation of expressions with free variables extended 
> with type casts

We disagree with this characterization.

Only the external type system comes from Hazelnut (Sec 3.1). That paper did 
not define a dynamic semantics at all.

Unlike gradually typed calculi, we do not abort on cast failure. Furthermore, 
we consider not only type holes but also expression holes, which operate not 
merely as free variables (which, as suggested, would be uninteresting), but 
as metavariable closures. Closure tracking is critical to the features in  
Sections 2 and 4.

Closure tracking has a relationship with CMTT (a non-obvious 
and satisfying logical connection, we believe!), but that work also did not 
define a dynamic semantics, nor did it consider non-empty holes or the 
commutativity property.

> calculus lacks recursion and control-flow

Please see above regarding recursion. The sum types extension in Appendix B 
adds control flow. We plan to move this into the main body.

> providing a good user interface for inspecting the resulting reified values 
> is the hard problem

Both theoretical and UI considerations are worthy of substantial 
investigation. This paper focuses on the former, motivated by a  
proof-of-concept UI design and several small but non-trivial examples.

This UI was designed with recursion in mind -- it allows you to  
to explore the recursive structure of closures iteratively (Sec 2.2). 
It does not display nested closures all at once, as in the 
formal notation. [more below the fold]

Branches blocked on a hole can certainly "blow up" in size. A partial 
solution is to visually collapse such subexpressions by default. Because we 
define a small step semantics, another approach is to pause evaluation when 
this occurs, continuing only if requested or, because we leave evaluation order
undefined, allowing the user to shift the "evaluation cursor" elsewhere.

>> effects and mutation

The interactions between gradual typing and mutation are well-studied and 
the standard approach (treating refs as invariant wrt consistency) should 
be applicable to recover the properties of Section 3. However, the 
commutativity theorem in Section 4 would not hold.

We anticipate that future work will explore variants of fill-and-resume
that support a weaker commutativity property (e.g. via checkpointing).

However, in order to provide the simplest and strongest theoretical
guarantees, we did not include mutation. (Note that Hazel is 
based on Elm, which is pure and quite practical.)
Thank you for your thoughtful feedback!

# Review A

> it doesn't even have fixpoint

Type holes actually give you general recursion (e.g. via the Y 
combinator) - see footnote 9 and [Siek and Taha, 2006] on embedding the 
untyped lambda calculus.

That said, we agree that an explicit fix operator might be more clear; if 
requested, we would be happy to make this small + simple addition.

> Are the screenshots in section 2 real, or only the ones in the appendix?

The appendix shows the submitted implementation, which 
supports all of the live programming features in Section 2 for 
the language of Section 3 plus Appendix B.

For the sake of exposition, the figures in Section 2 do go beyond the language
of Section 3 (acknowledged in the intro to Section 2). To create these figures, 
we ran simpler variants of the examples (using typical lambda encodings), then 
manually edited the HTML/CSS only to add conveniences from Elm, namely record 
labels, pattern matching and sugar for functions and lists.

(We plan to submit a more recent snapshot of Hazel that substantially narrows 
or eliminates this small "convenience gap" for artifact evaluation.)

# Review C

> mostly based prior work (Hazelnut type system), or a variation on gradually 
> typed calculi, and not very surprising. In fact, evaluation of expression 
> with holes is akin to evaluation of expressions with free variables extended 
> with type casts

We disagree with this characterization.

Only the external type system comes from Hazelnut (Sec 3.1). That paper did 
not define a dynamic semantics at all.

Unlike gradually typed calculi, we do not abort on cast failure. Furthermore, 
we consider not only type holes but also expression holes, which operate not 
merely as free variables (which, as suggested, would be uninteresting), but 
as metavariable closures. Closure tracking is critical to the features in  
Sections 2 and 4.

Closure tracking has a relationship with CMTT (a non-obvious 
and satisfying logical connection, we believe!), but that work also did not 
define a dynamic semantics, nor did it consider non-empty holes or the 
commutativity property.

> calculus lacks recursion and control-flow

Please see above regarding recursion. The sum types extension in Appendix B 
adds control flow. We plan to move this into the main body.

> providing a good user interface for inspecting the resulting reified values 
> is the hard problem

Both theoretical and UI considerations are worthy of substantial 
investigation. This paper focuses on the former, motivated by a  
proof-of-concept UI design and several small but non-trivial examples.

This UI was designed with recursion in mind -- it allows you to  
to explore the recursive structure of closures iteratively (Sec 2.2). 
It does not display nested closures all at once, as in the 
formal notation. [more below the fold]

Branches blocked on a hole can certainly "blow up" in size. A partial 
solution is to visually collapse such subexpressions by default. Because we 
define a small step semantics, another approach is to pause evaluation when 
this occurs, continuing only if requested or, because we leave evaluation order
undefined, allowing the user to shift the "evaluation cursor" elsewhere.

>> effects and mutation

The interactions between gradual typing and mutation are well-studied and 
the standard approach (treating refs as invariant wrt consistency) should 
be applicable to recover the properties of Section 3. However, the 
commutativity theorem in Section 4 would not hold.

We anticipate that future work will explore variants of fill-and-resume
that support a weaker commutativity property (e.g. via checkpointing).

However, in order to provide the simplest and strongest theoretical
guarantees, we did not include mutation. (Note that Hazel is 
based on Elm, which is pure and quite practical.)

-------

Regarding , the UI displays a "breadcrumb" path from the result to
reflect the exploration that the user has done (the basic UI can be seen at the
bottom of Fig. 17). We would be happy to include this more prominently in the
second phase submission (it requires showing a second, more deeply nested hole
environment, so we will need the additional space available in the second phase
submission).