### 2018-07-07 examples.tex

% !TEX root = hazelnut-dynamics.tex
%% \newcommand{\examplesSec}{Live Functional Programming in Hazel, By Example}
%% \newcommand{\examplesSec}{Live Programming in \HazelnutLive, By Example}
\newcommand{\examplesSec}{Live Functional Programming with Typed Holes in Hazel}
\section{\protect\examplesSec} % don't like the all-caps thing that the template does, so protecting it from that
\label{sec:examples}

\input{fig-grades-example}

\newcommand{\overviewExample}[2]{\paragraph{Example {#1}: {#2}}}

This section gives an example-driven overview of our  approach as implemented in  \Hazel, a browser-based lab notebook environment being developed by \citet{HazelnutSNAPL}. The \Hazel user interface is  based roughly on IPython/Jupyter \cite{PER-GRA:2007} and the \Hazel language is tracking toward feature parity with \Elm~(\url{elm-lang.org}) \cite{czaplicki2012elm,Elm}, a popular pure functional programming language similar to ``core ML'', with which we assume basic  familiarity.
For the sake of 
exposition, we have ``post-processed'' the screenshots in this section to make use of a few language conveniences from \Elm~that were not available in \Hazel as of this writing, e.g. pattern matching in function arguments, list  notation, and record labels (currently there are only tuples). These conveniences are entirely orthogonal to the contributions of this paper: all of the live programming features shown in this section have been implemented, both in \Hazel itself and in the simplified companion implementation of the core calculus included in the supplement (see Sec.~\ref{sec:implementation}).

\input{examples-1}
\input{examples-2}
\input{examples-3}
\input{examples-4}

% \paragraph{Recap}
% %
% In the above four subsections, respectively, we considered programs with:
% %
% (1) empty expression holes,
% %
% (2) non-empty expression holes for ill-typed expressions,
% %
% (3) type holes, and
% %
% (4) non-empty expression holes for well-typed expressions.
% %
% Next, we will formally describe the novel dynamic semantics of \HazelnutLive{}
% that allows programs with combinations of these kinds of incompleteness to be
% evaluated.